---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  fig.width = 10,
  fig.height = 10
)
```

# ESDist: Calculate and visualise field-specific ESDs

<!-- badges: start -->
<!-- badges: end -->

## Description
The `ESDist` package is designed to calculate and visualise field-specific
effect size distributions, based on data that can easily be obtained from 
meta-analyses.


### Installation

You can install the ESDist package from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("berntgl/ESDist")
```

## esd_plot()
### Plot a simple ESD

The `esd_plot()` function can visualise ESDs based on effect size estimates 
obtained from meta-analyses.

```{r, message=FALSE, warning=FALSE}
library(ESDist)
library(dplyr)
library(ggplot2)
```

```{r}
plot1 <- esd_plot(df = ot_dat,
         es = yi,
         es_type = "Hedges' g")

plot1
```


By adding `mean = "mean"`, we can plot the true mean of this distribution.

```{r}
plot2 <- esd_plot(df = ot_dat,
         es = yi,
         es_type = "Hedges' g",
         mean = "mean")

plot2
```


### Plot effect size benchmarks

It is also possible to plot effect size benchmarks based on the 25th, 50th, and 
75th percentiles by adding `method = "quads"` (or based on the 16.65th, 50thf, 
and 83.35th percentiles by adding `method = "thirds"`). Note that only absolute 
effect sizes are used to accomplish this.

```{r}
plot3 <- esd_plot(df = ot_dat,
         es = yi_abs,
         es_type = "Hedges' g",
         method = "quads")

plot3
```


### Plot empirical effect size ranges

Finally, we can specify the range of effect sizes that is equal to or larger 
than a specified `sesoi`. Note that only absolute effect sizes are used to 
accomplish this.

```{r}
plot4 <- esd_plot(df = ot_dat,
         es = yi_abs,
         es_type = "Hedges' g",
         sesoi = 0.2)

plot4
```


## esd_plot_group()
### Simple plot per group

The `esd_plot_group()` function allows for specifying a `grouping_var` to group 
data and create plots for each group with 20 or more effect sizes. 

```{r}
plot5 <- esd_plot_group(df = ot_dat,
         es = yi,
         es_type = "Hedges' g",
         grouping_var = group)

plot5
```


### Plot effect size benchmarks per group.
Like the `esd_plot()` function, the `esd_plot_group()` function lets you plot
effect size benchmarks based on a specific set of percentiles, by adding 
`method = "quads"` or `method = "thirds"`. Note that only absolute effect sizes
are used to accomplish this.

```{r}
plot6 <- esd_plot_group(df = ot_dat,
         es = yi_abs,
         es_type = "Hedges' g",
         grouping_var = group,
         method = "thirds")

plot6
```


## esd_table()
### Calculate effect size benchmarks for whole dataset
By using the `esd_table()` function, you can calculate the effect size 
benchmarks for your dataset by specifying the dataset (`df`), and the column 
containing all effect sizes (`es`). This will give you the effect size 
benchmarks based on the 25th, 50th, and 75th percentiles. Note that only 
absolute effect sizes should be used.

```{r, message=FALSE, warning=FALSE}
library(ESDist)
library(dplyr)
```

```{r}
table1 <- esd_table(df = ot_dat,
                    es = yi_abs)

table1
```

Alternatively, you can calculate benchmarks by splitting the distribution into 
three equal-sized chunks, and calculating the median of every third, which would
correspond to the 16.65th, 50th, and 83.35th percentiles. This can be achieved
by adding `method = "thirds"` to the code.

```{r}
table2 <- esd_table(df = ot_dat,
                    es = yi_abs,
                    method = "thirds")

table2
```

### Calculate effect size benchmarks per group
By specifying `grouping_var`, the user can calculate the effect size benchmarks
per group, for every group with at least three effect sizes. (The number of 
required effect sizes per group can be overwritten by specifying 
`min_group_size`.)

```{r}
table3 <- esd_table(df = ot_dat,
                    es = yi_abs,
                    grouping_var = group)

table3
```

### Editing and saving tables
The values presented in the table default to two decimal places. This can be 
overwritten by setting `ndec` to the desired number of decimal places. 
Tables can be saved by setting `csv_write` to `TRUE`. Optionally, by setting
`file_name` to a string specifying the sub-directory and file name (e.g., 
`sub-folder/table.csv`), the table can be saved to a specific location with a
specific name. 

## esd_perc()
### Calculating the percentile of a value
Using `esd_perc()`, you can calculate to which percentile of the ESD a given 
value corresponds. Note that absolute effect size values should be used.

```{r, message=FALSE, warning=FALSE}
library(ESDist)
library(dplyr)
library(ggplot2)
```

```{r}
esd_perc(df = ot_dat,
         es = yi_abs,
         value = 0.2)
```
