---
title: "ESDist: Calculate and visualise field-specific ESDs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ESDist: Calculate and visualise field-specific ESDs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 10,
  fig.height = 7
)
```

# ESDist: Calculate and visualise field-specific ESDs

<!-- badges: start -->
<!-- badges: end -->

## Description
The `ESDist` package is designed to calculate and visualise field-specific
effect size distributions, based on data that can easily be obtained from 
meta-analyses.

[![DOI](https://zenodo.org/badge/532824630.svg)](https://zenodo.org/badge/latestdoi/532824630)


### Installation

You can install the `ESDist` package from [GitHub](https://github.com/) with:

``` {r}
# install.packages("devtools")
library(devtools)
devtools::install_github("berntgl/ESDist")
```
```{r, include=FALSE}
devtools::load_all()
```


### Data structure

The `ESDist` package is designed for datasets containing data that can easily 
be extracted from pre-existing meta-analyses. As such, most functions in this 
package work on a dataset that only contains a column with effect sizes. 
However, for some functions and additional functionalities, users might want to 
include some grouping variable or other study information. For the publication
bias-adjusted functions, users need a column with the standard error for each
study effect size. Alternatively, users can include a column with the lower 
bound of the 95% confidence interval, and one for the upper bound. The `ot_dat` 
dataset included in the `ESDist` package contains several columns in adition to 
the effect size column (`yi`).

```{r}
head(ESDist::ot_dat)
```


## esd_plot()
### Plot a simple ESD

The `esd_plot()` function can visualise ESDs based on effect size estimates 
obtained from meta-analyses.

```{r, message=FALSE, warning=FALSE}
library(ESDist)
```

```{r, fig.height = 7}
plot1 <- esd_plot(df = ot_dat,
         es = yi,
         es_type = "Hedges' g")

plot1
```


### Plot effect size benchmarks

It is also possible to plot effect size benchmarks based on the 25th, 50th, and 
75th percentiles by adding `method = "quads"` (or based on the 16.65th, 50th, 
and 83.35th percentiles by adding `method = "thirds"`).

```{r, fig.height = 7}
plot2 <- esd_plot(df = ot_dat,
         es = yi,
         es_type = "Hedges' g",
         method = "quads")

plot2
```


### Plot empirical effect size ranges

Finally, we can specify the range of effect sizes that is equal to or larger 
than a specified `sesoi`.

```{r, fig.height = 7}
plot3 <- esd_plot(df = ot_dat,
         es = yi,
         es_type = "Hedges' g",
         sesoi = 0.3)

plot3
```


## esd_plot_group()
### Simple plot per group

The `esd_plot_group()` function allows for specifying a `grouping_var` to group 
data and create plots for each group with 20 or more effect sizes. 

```{r, fig.height = 7}
plot4 <- esd_plot_group(df = ot_dat,
         es = yi,
         es_type = "Hedges' g",
         grouping_var = group)

plot4
```


### Plot effect size benchmarks per group.
Like the `esd_plot()` function, the `esd_plot_group()` function lets you plot
effect size benchmarks based on a specific set of percentiles, by adding 
`method = "quads"` or `method = "thirds"`.

```{r, fig.height = 7}
plot5 <- esd_plot_group(df = ot_dat,
         es = yi,
         es_type = "Hedges' g",
         grouping_var = group,
         method = "quads")

plot5
```

## esd_plot_pba()
### Plot an adjusted ESD against a raw ESD.
Using the `meta` and `metasens` packages (Schwarzer et al., 2023), we can 
adjust every individual effect size for publication bias, and plot the adjusted 
distribution against theraw distribution. 

First, we need to import the `meta` and `metasens` packages

```{r, message=FALSE, warning=FALSE}
library(meta)
library(metasens)
```

Next, we create an object of class 'meta' (m1), and subsequently use this
to create an object of class 'limitmeta' (l1). 

```{r, message=FALSE, warning=FALSE}
m1 <- metagen(TE = ot_dat$yi, seTE = ot_dat$sei)
l1 <- limitmeta(m1)
```

We can now use the `l1` object to create our distribution.

```{r, fig.height = 7}
plot6 <- esd_plot_pba(lim_obj = l1, 
          es_type = "Hedges' g")

plot6
```

Alternatively, we can plot the effect size benchmarks for both distributions.

```{r, fig.height = 7}
plot7 <- esd_plot_pba(lim_obj = l1, 
          es_type = "Hedges' g",
          method = "quads")
plot7
```

Or the range of effect sizes larger than or equal to a SESOI

```{r, fig.height = 7}
plot8 <- esd_plot_pba(l1, 
          "Hedges' g",
          sesoi = 0.3)
plot8
```

## esd_table()
### Calculate effect size benchmarks for whole dataset
By using the `esd_table()` function, you can calculate the effect size 
benchmarks for your dataset by specifying the dataset (`df`), and the column 
containing all effect sizes (`es`). This will give you the effect size 
benchmarks based on the 25th, 50th, and 75th percentiles.

```{r, message=FALSE, warning=FALSE}
library(ESDist)
library(dplyr)
```

```{r}
table1 <- esd_table(df = ot_dat,
                    es = yi)

table1
```

### Calculate effect size benchmarks per group
By specifying `grouping_var`, the user can calculate the effect size benchmarks
per group, for every group with at least three effect sizes. (The number of 
required effect sizes per group can be overwritten by specifying 
`min_group_size`.)

```{r}
table3 <- esd_table(df = ot_dat,
                    es = yi,
                    grouping_var = group)

table3
```


## esd_table_pba()
The `esd_table_pba()` function allows users to use an object of type `limitmeta`
to calculate benchmarks based on effect sizes that are adjusted for publication
bias. 

```{r}
table4 <- esd_table_pba(lim_obj = l1)

table4
```

In case the user wants effect size benchmarks per group, the user should define
the `subgroup` argument when defining the `meta` object:

```{r, message=FALSE, warning=FALSE}
m2 <- metagen(TE = ot_dat$yi, seTE = ot_dat$sei, subgroup = ot_dat$group)
l2 <- limitmeta(m2)
```

Then, in the `esd_table_pba()` function, the `grouping` argument should be set
to `TRUE`. 

```{r}
table5 <- esd_table_pba(lim_obj = l2,
                    grouping = TRUE)

table5
```

### Editing and saving tables
The values presented in the table default to two decimal places. This can be 
overwritten by setting `ndec` to the desired number of decimal places. 
Tables can be saved by setting `csv_write` to `TRUE`. Optionally, by setting
`file_name` to a string specifying the sub-directory and file name (e.g., 
`sub-folder/table.csv`), the table can be saved to a specific location with a
specific name. 

## esd_perc()
### Calculating the percentile of a value
Using `esd_perc()`, you can calculate to which percentile of the ESD a given 
value corresponds. Note that absolute effect size values should be used.

```{r, message=FALSE, warning=FALSE}
library(ESDist)
library(dplyr)
library(ggplot2)
```

```{r}
esd_perc(df = ot_dat,
         es = yi,
         value = 0.3)
```

